Improving Emotion Detection through Real-Time Translation of Text to ML Models Trained in Different Languages

### 2023-08-04 paper/main.tex
Paper Updates

\tableofcontents
\clearpage

\section{Introduction}

@@

\section{Literature Review}
To ensure comprehensive coverage of the latest developments and research in ED for this research paper, consideration and review was only given to incorporating those research papers and online news articles that were published after the year 2018, thereby providing an up-to-date perspective on the subject matter.

\subsection{Emotion Detection Importance and Diverse Populations}
Emotion Detection (ED) holds significant importance in various domains, including psychology, social sciences, and human-computer interaction. It enables the analysis of human emotions at scale, providing valuable insights into individual and collective emotional states.

@@

\section{Methodology}

@@

Each translated file now contains four (4) columns that are labeled as in Tables \ref{table:pd_de_translated} and \ref{table:pd_en_translated} with the exact same column naming.

@@

 Both sentences are then processed in real-time on both the English and German models for the Emotion Detection and will be compared to each other's output. The average of the prediction will be used for final prediction of the supplied sentence and the emotion that was detected.

 \begin{table}[h!]
\centering
\begin{tabular}{ | L | L | L | L | }
    \hline
    \multicolumn{4}{|c|}{ML Models } \\
    \hline
    Model & 
    Data &
    Rows for Training & 
    Rows for Testing  \\
    \hline
    A &
    English Original &
    1200 & 
    300 (20\%) \\
    \hline
    B &
    English Extended By German &
    2700 & 
    300 same as Model "A" \\
    \hline
    C &
    German Original &
    1200 & 
    300 (20\%) \\
    \hline
    D &
    German Extended By English &
    2700 & 
    300 same as Model "C" \\

@@

\subsection{Evaluation Metrics and methodologies for measuring the performance of the ML models}
The evaluation of ML models trained on text datasets is a critical step in assessing their performance and reliability. Our research project is no different and the following assumptions are used to determine the effectiveness of the methodology of extending our datasets:
\begin{itemize}
\item \textbf{Accuracy:} Accuracy measures the overall correctness of the emotion predictions made by the model. It is calculated as the ratio of correctly classified instances to the total number of instances in the dataset. However, accuracy may not provide a complete picture, especially if the dataset is imbalanced.
\item \textbf{Precision, Recall, and F1 Score:} Precision represents the proportion of correctly predicted emotions among all instances predicted as a specific emotion. Recall measures the proportion of correctly predicted emotions among all instances that truly belong to a specific emotion. F1 score combines precision and recall into a single metric, providing a balanced evaluation.
\item \textbf{Cross-Validation:} Cross-validation techniques, such as k-fold or stratified cross-validation, divide the dataset into multiple subsets for training and testing. This helps evaluate the model's performance across different data samples, mitigating over-fitting and providing a more robust estimation of its generalization ability.
\item \textbf{Error Analysis:} Performing error analysis on the model's predictions helps identify specific patterns or instances where the model might struggle or make frequent mistakes. This analysis can guide improvements and provide insights for future model development.
\end{itemize}

@@

\subsubsection{Written \& Oral Presentation of Research Finding}
In conclusion of the project a research paper and oral presentation is required to pass MTSU Computation Science PhD Qualifying Exam. Such paper is being presented with this paper and the oral presentation will be in Microsoft Power Point format.

The listing the datasets used, code for the application, and details of the research project process in this research paper is openly available and hosted on a GitHub repository, allowing for easy access, reproducible, and collaborative engagement with the research findings. The repository is open and available under: \texttt{https://github.com/richardhoehn/mtsu.coms.qualifying-exam}\cite{Hoehn_Improving_Emotion_Detection_2023}

\section{Results and Analysis}

\subsection{Presentation of the prediction results obtained from the extended datasets}

@@

\subsection{Comparison of Prediction Result of the Original and Extended Language (translated) Datasets}
\ldots

@@

\section{Demonstration and Application}

>>>>

\tableofcontents

\clearpage
\section{Introduction}

@@

\clearpage
\section{Literature Review}
To ensure comprehensive coverage of the latest developments and research in ED for this research paper, consideration and review was only given to incorporating those research papers and online news articles that were published after the year 2018, thereby providing an up-to-date perspective on the subject matter.

\subsection{Emotion Detection Importance and Diverse Populations}
Emotion Detection (ED) holds significant importance in various domains, including psychology, social sciences, and human-computer interaction. It enables the analysis of human emotions at scale, providing valuable insights into individual and collective emotional states. In Chowanda et. al. paper\cite{CHOWANDA-2021821} they believe that "Emotions hold a paramount role in the conversation, as it expresses context to the conversation."

@@

\clearpage
\section{Methodology}

@@

Each translated file now contains four (4) columns that are labeled in Tables \ref{table:pd_de_translated} and \ref{table:pd_en_translated} with the exact same column naming.

@@

 Using the datasets from Tables \ref{table:pd_de_translated}, \ref{table:pd_en_translated}, and \ref{table:ml_models} that list an approximate $\sim$1,275 and $\sim$2,775 features for training. The algorithms tested in this research project ranged from the traditional machine learning to deep learning techniques, they are Na\"ive Bayes, Generalised Linear Model (GLM), and Random Forest Classifier models. Each of these algorithms used and their results are listed in the Results and Analysis Section \ref{sec:results-and-analysis} below. In Chowanda et al. paper\cite{CHOWANDA-2021821} on Text-based Emotion Techniques multiple algorithms were proposed, this research focused on just three out of their six since the others had lower prediction rates based on Chowanda et al. paper\cite{CHOWANDA-2021821} paper.

 \begin{table}[h!]
\centering
\begin{tabular}{ | L | L | L | L | }
    \hline
    \multicolumn{4}{|c|}{ML Models } \\
    \hline
    Model & 
    Data &
    Rows for Training (85\%) & 
    Rows for Testing (15\%)  \\
    \hline
    A &
    English Original &
    1,275 & 
    225 \\
    \hline
    B &
    English Extended By German &
    2,775 & 
    225 same as Model "A" \\
    \hline
    C &
    German Original &
    1,275 & 
    225 \\
    \hline
    D &
    German Extended By English &
    2,775 & 
    225 same as Model "C" \\

@@

\subsection{Evaluation Metrics and methodologies for measuring the performance of the ML models}
The evaluation of ML models trained on the English, German and their respective extension datasets is a critical step in assessing their performance and reliability. We chose to use the MulticlassClassificationEvaluator\footnote{URL: https://spark.apache.org/docs/latest/api/python/reference/api/pyspark.ml.evaluation.MulticlassClassificationEvaluator.html} because is a well known tool in the field of natural language processing (NLP) research, specifically in multi class classification tasks such as Emotion Detection (ED). Generally this evaluator is used to evaluate the performance of machine learning (ML) models, using transformer-based architectures like BERT, RoBERTa, or DistilBERT\cite{Sanh2019-DistilBERT-AD}, with DistilBERT\cite{Sanh2019-DistilBERT-AD} being the one use in this research.

The MulticlassClassificationEvaluator provides the essential evaluation metrics, including accuracy, precision, recall, and F1 score, for ED classifications. These metrics provide valuable insights into a model's ability to correctly classify emotions across multiple classes and language identification. Our research project is no different and the following that the assumptions are used to determine the effectiveness of the methodology of extending our datasets:
\begin{itemize}

\item \textbf{Accuracy:} Accuracy measures the overall correctness of the emotion predictions made by the model. It is calculated as the ratio of correctly classified instances to the total number of instances in the dataset. However, accuracy may not provide a complete picture, especially if the dataset is imbalanced.

\item \textbf{Precision, Recall, and F1 Score:} Precision represents the proportion of correctly predicted emotions among all instances predicted as a specific emotion. Recall measures the proportion of correctly predicted emotions among all instances that truly belong to a specific emotion. F1 Score ($F_{1}=\frac{\text{Precision} \times \text{Recall}}{\text{Precision} + \text{Recall}}$) combines precision and recall into a single metric, providing a balanced evaluation.

\item \textbf{Error Analysis:} Performing error analysis on the model's predictions helps identify specific patterns or instances where the model might struggle or make frequent mistakes. This analysis can guide improvements and provide insights for future model development.
\end{itemize}

@@

\subsection{Written \& Oral Presentation of Research Finding}
In conclusion of the project a research paper and oral presentation is required to pass MTSU Computation Science PhD Qualifying Exam. The listing the datasets used, code for the application, and details of the research project process in this research paper is openly available and hosted on a GitHub repository, allowing for easy access, reproducible, and collaborative engagement with the research findings. The repository is open and available under: \texttt{https://github.com/richardhoehn/mtsu.coms.qualifying-exam}\cite{Hoehn_Improving_Emotion_Detection_2023}

\clearpage
\section{Results and Analysis}
\label{sec:results-and-analysis}
The following section lists the prediction rates of training and testing of the original and extended datasets both in English and German. Overall results seem to indicate that there was no and even negative results based on adding the extended data for Emotion Detection. In the below sub sections details on the results are listed in prose, tables, and formauls.

\subsection{Prediction Results of Original \& Extended Datasets}

@@

\subsection{Comparison of Prediction Rates of Original and Extended Language Datasets}
\ldots

@@

\subsection{Conclusion and Future Work}
\ldots

\clearpage
\section{Demonstration and Application}
